---
title: "R Assignment 2"
author: "Jared Barber"
date: "2/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 4.2
a. Store the vector c(7,1,7,10,5,9,10,3,10,8) as foo. Identify the elements greater than 5 OR equal to 2.
``` {r 4.2.a}
foo<-c(7,1,7,10,5,9,10,3,10,8)
foo > 5 | foo == 2 
```
b. Store the vector c(8,8,4,4,5,1,5,6,6,8) as bar. Identify the ele-
ments less than or equal to 6 AND not equal to 4.
```{r 4.2.b}
bar<-c(8,8,4,4,5,1,5,6,6,8)
bar <= 6 & bar != 4
```
c. Identify the elements that satisfy (a) in foo AND satisfy (b) in bar.
```{r 4.2.c}
(foo > 5 | foo == 2) & (bar <= 6 & bar!=4)
```
d. Store a third vector called baz that is equal to the element-wise sum of foo and bar. Determine the following:
i. The elements of baz greater than or equal to 14 but not equal to 15 
ii. The elements of the vector obtained via an element-wise division of baz by foo that are greater than 4 OR less than or equal to 2
```{r 4.2.d}
baz = foo + bar
baz >= 14 & baz != 15
t = baz / foo
t > 4 | t <= 2
```
## Exercise 4.3
a.Store this vector of 10 values: foo <- c(7,5,6,1,2,10,8,3,8,2).
Then, do the following:
i. Extract the elements greater than or equal to 5, storing the result as bar.
ii. Display the vector containing those elements from foo that remain after omitting all elements that are greater than or equal to 5.
```{r 4.3.a}
foo<-c(7,5,6,1,2,10,8,3,8,2)
#i
bar<- foo[foo >= 5]
bar
#ii
foo[foo<5]
```
b. Use bar from (a)(i) to construct a 2 × 3 matrix called baz, filled in a row-wise fashion. Then, do the following:
i. Replace any elements that are equal to 8 with the squared value of the element in row 1, column 2 of baz itself.
ii. Confirm that all values in baz are now less than or equal to 25 AND greater than 4.
```{r 4.3.b}
baz<-matrix(bar,2,3,byrow = T)
#i
baz[baz==8]<-baz[1,2]^2
#ii
baz<=25&baz>4
```
## Exercise 4.5
a. Use your knowledge of vectors (for example, subsetting and overwriting) to create two character vectors: sex with entries "M" (male) and "F" (female) and party with entries "National", "Labour", "Greens", "Maori", and "Other". Make sure the entries are placed in the correct positions as outlined earlier.
```{r 4.5.a}
sex<-rep('M', 20)
sex[c(1,5:7,12,14:16)] = 'F'

party<-rep('National', 20)
party[c(1,4,12,15,16,19)] = 'Labour'
party[c(6,9,11)] = 'Greens'
party[c(10,20)] = 'Other'

sex
party
```
b. Create two different factor vectors based on sex and party. Does it make any sense to use ordered=TRUE in either case? How has R appeared to arrange the levels?
```{r 4.5.b}
sex<-factor(sex)
sex
party<-factor(party)
party
```
No, there is no reason to use 'ordered=TRUE' in either case. R has arranged the levels lexically, which is not meaningful. 

c. Use factor subsetting to do the following:
i. Return the factor vector of chosen parties for only the male participants.
ii. Return the factor vector of genders for those who chose National.
```{r 4.5.c}
#i
t<-factor(party[c(which(sex=='M'))])
t
#ii
t<-factor(sex[c(which(party=='National'))])
t
```
d. Another six people joined the survey, with the results c("National","Maori","Maori","Labour","Greens","Labour") for the preferred party and c("M","M","F","F","F","M") as their gender. Combine these results with the original factors from (b).Suppose you also asked all individuals to state how confident they were that Labour will win more seats in Parliament than National in the next election and to attach a subjective percentage to that confidence. The following 26 results were obtained: 93, 55, 29, 100, 52, 84, 56, 0, 33, 52, 35, 53, 55, 46, 40, 40, 56, 45, 64, 31, 10, 29, 40, 95, 18, 61.
```{r 4.5.d}
p<-factor(c("National","Maori","Maori","Labour","Greens","Labour"))
c(party,p)
s<-factor(c("M","M","F","F","F","M"))
c(sex, s)
```
e. Create a factor with levels of confidence as follows: 
Low for percentages [0,30];
Moderate for percentages (30,70]; and
High for percentages (70,100].
```{r 4.5.e}
conf <-c(93, 55, 29, 100,52, 84, 56, 0, 33, 52, 35, 53, 55, 46, 40, 40, 56, 45, 64, 31, 10, 29, 40, 95, 18, 61)
conf[(c(which(conf >70)))] = 'High'
conf[c(which(conf<=30))]= 'Low'
conf[c(which(conf<=70))]= 'Moderate'

conf<-factor(conf)
conf

```
f. From (e), extract the levels corresponding to those individuals who originally said they identified with Labour. Do this also for National. What do you notice?
```{r 4.5.f}
t<-factor(conf[c(which(party=='Labour'))])
t
t<-factor(conf[c(which(party=='National'))])
t
```
Those individuals who identified with Labour are more confident Labour will win more seats than National in the upcoming election. The individuals who identified with National are less confident.

## Exercise 9.2
a. Use positional matching with seq to create a sequence of values between −4 and 4 that progresses in steps of 0.2.
```{r 9.2.a}
vec<-c(seq(-4,4,0.2))
vec
```
b. In each of the following lines of code, identify which style of argument matching is being used: exact, partial, positional, or mixed. If mixed, identify which arguments are specified in each style.
ii. rep(1:2,3)
iii. seq(from=10,to=8,length=5)
iv. sort(decreasing=T,x=c(2,1,1,2,0.3,3,1.3))
v. which(matrix(c(T,F,T,T),2,2))
vi. which(matrix(c(T,F,T,T),2,2),a=T)

ii. Positional
iii. Mixed - 'from' and 'to' are exact, 'length' is partial
iv. Exact
v. Positional
vi. Mixed - 'x','data','nrow', and 'ncol' are positional, 'arr.ind' is partial.

## Exercise 16.1
b. What is the probability you see birds at all sites? This is a binomial probability. P(x = 13) with n = 13 and p = 0.75 is 
```{r 16.1.b}
dbinom(13,13,0.75)
```
c. What is the probability you see birds at more than 9 platforms?
P(X > 9) = 1-P(x<=9)
```{r 16.1.c}
1-pbinom(9,13,0.75)
```
d. What is the probability of seeing birds at between 8 and 11 platforms (inclusive)? Confirm your answer by using only the d-function and then again using only the p-function.
P(8<=x<=13) = Σ p(x=8), p(x=9), p(x=10),p(x=11)
P(8<=x<=13) = P(x<=13)-P(x<8) = P(x<=13)-p(x<=7)
```{r 16.1.d}
dbinom(8,13,0.75)+dbinom(9,13,0.75)+dbinom(10,13,0.75)+dbinom(11,13,0.75)
pbinom(11,13,0.75)-pbinom(7,13,0.75)
```
e. Say that, before your visit, you decide that if you see birds at fewer than 9 sites, you’ll make a scene and demand your entry fee back. What’s the probability of your embarrassing yourself in this way?
P(x<9) = P(x<=8)
```{r 16.1.e}
pbinom(8,13,0.75)
```
## Exercise 16.2
a. What is the probability that more than 100 cars pass her on any given Saturday?

``` {r 16.2.a}
1-ppois(100,107)
```
b. Determine the probability that no cars pass
```{r 16.2.b}
dpois(0,107)
```
## Exercise 16.3
You visit a national park and are informed that the height of a certain
species of tree found in the forest is uniformly distributed between 3
and 70 feet.
a. What is the probability you encounter a tree shorter than 5 1/2 feet?
```{r 16.3.a}
punif(5.5,3,70)
```
b. For this probability density function, what is the height that marks the cutoff point of the tallest 15 percent of trees?
```{r 16.3.b}
#b Cutoff points of tallest 15 percent of trees.
qunif((1-0.15),3,70)
```
## Exercise 16.4
a. A tutor knows that the length of time taken to complete a certain statistics question by first-year undergraduate students, X, is normally distributed with a mean of 17 minutes and a standard deviation of 4.5 minutes.
i. What is the probability a randomly selected undergraduate takes more than 20 minutes to complete the question?
```{r 16.4.a.i}
1-pnorm(20,17,4.5)
```
ii. What’s the chance that a student takes between 5 and 10 minutes to finish the question?
``` {r 16.4.a.ii}
pnorm(10,17,4.5)-pnorm(5,17,4.5)
```
iii. Find the time that marks off the slowest 10 percent of students
```{r 16.4.a.iii}
qnorm((1-0.10),17,4.5)
```
b. A meticulous gardener is interested in the length of blades of grass on his lawn. He believes that blade length X follows a normal distribution centered on 10 mm with a variance of 2 mm.
i. Find the probability that a blade of grass is between 9.5 and 11 mm long.
```{r 16.4.b.i}
pnorm(11,10,sqrt(2))-pnorm(9.5,10,sqrt(2))
```
ii. What are the standardized values of 9.5 and 11 in the context of this distribution? Using the standardized values, confirm that you can obtain the same probability you found in (i) with the standard normal density.
```{r 16.4.b.ii}
#ii

low = (9.5-10)/sqrt(2)
high= (11-10)/sqrt(2)

low
high

pnorm(high,0,1)-pnorm(low,0,1)
```
iii. Below which value are the shortest 2.5 percent of blade lengths found?
```{r 16.4.b.iii}
qnorm(0.025,10,sqrt(2))
```
iv. Standardize your answer from (iii).
```{r 16.4.b.iv}
qnorm(0.025,0,1)
```




